---
title: "WorldClim"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = F)
library(tidyverse)
library(raster)
```

Get the worldclim2.0 data at high spatial resolution (30s) for all accessions.


```{r Data}
clim1 <- read_tsv('/Volumes/nordborg/user/pieter.clauw/Documents/Source/Accessions/List_2029_latlon_env.txt')
araclim <- read_csv('/Volumes/nordborg/user/pieter.clauw/Documents/Source/Accessions/araclim.csv')

# accesisons
acns.modif <- read_csv('/Volumes/nordborg/user/pieter.clauw/Documents/Source/Accessions/accs_2029_ids_lat_long.modified_MN_SH.csv')

# default coordinate system
crs.wc <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'

clim.dir <- '/Volumes/nordborg/user/pieter.clauw/Documents/Source/WorldClim/wc2.0_30s/'
clim.vars <- c('bio', 'prec', 'srad', 'tavg', 'tmax', 'tmin', 'vapr', 'wind')
```
Wordlclim data uses the longlat WGS84 coordinate system (CRS).This is the global system also use by google maps and the same as the coordinates for the accessions.

coordinates in List_2029_latlon_env are not matching the master list.
e.g. the deviation for the coordinates of accession 6074 differ between land and sea with ±500m.

```{r get data per accession}
clim2 <- acns.modif %>%
  dplyr::select(accession = tg_ecotypeid, longitude, latitude)

for (cv in clim.vars)
{
  cv.files <- list.files(paste(clim.dir, cv, sep = ''), pattern = '*.tif', full.names = T)
  for (cv.file in cv.files)
  {
    # load data and test CRS
    cv.data <- raster(cv.file)
    if (crs(cv.data, asText = T) != crs.wc){print('WARNING: CRS is not correct')}
    
    cv_nr <- strsplit(strsplit(cv.file, split = 'wc2.0_30s_')[[1]][2], split = '[.]')[[1]][1]
    clim2[, cv_nr] <- extract(cv.data, clim2[, c('longitude', 'latitude')], method = 'bilinear')
  }
}
```


```{r test against clim1 and araclim}
# clim1
clim.vars.common <- colnames(clim2)[colnames(clim2) %in% colnames(clim1)]

clim.join <- clim2 %>%
  left_join(., clim1, by = c('accession' = 'tg_ecotypeid')) %>%
  dplyr::select(matches(".*x$|.*y$"))

test <- clim.join %>%
  dplyr::select(contains('tmin_10'))

plot(test$tmin_10.x, test$tmin_10.y)
```

Adding summaries per season.
Winter is jan, feb, mar
Spring is apr, may, jun
Summer is jul, aug, sep
Autumn is oct, nov, dec

Seasona are annotated with month numbers
  e.g. summer minimum temperature is tmin_07_08_09.
```{r make seasonal}

for (cv in clim.vars[-1])
{
  for (i in c(1,4,7,10))
  {
    months.nr <- sprintf("%02d", c(i:(i+2)))
    months.cv <- paste(cv, months.nr, sep = '_')
    season <- paste(cv, paste(months.nr, collapse = '_'), sep = '_')

    clim2 <- mutate(clim2, !!season := rowMeans(dplyr::select(clim2, months.cv)))
  }
}
```



```{r write data}
write_csv(clim2, '/Volumes/nordborg/user/pieter.clauw/Documents/Source/Accessions/2029_modified_MN_SH_wc2.0_30s_bilinear.csv')
```





