---
title: "AnnotateRawData"
author: "Pieter Clauw"
date: "11/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = F)
source('Rfunctions.R')
```

```{r data}
# raw lemnatech data
rep1 <- read.table('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth/RawData/Pieter_SepttoDez_2017_RawData.csv', header = T, sep = ';')
rep2_3 <- read.table('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth/RawData/LemnaTecResults_16v62018_2017Analysis.csv', header = T, sep = ';')

# hesoID annotations with positions
heso.rep1 <- read.table('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth/RawData/hesoAnnotations_SepttoDez_2017.txt', header = T, sep = '\t')
heso.rep2_3 <- read.table('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth/RawData/hesoAnnotations_MartoJune_2018_correct.txt', header = T, sep = '\t')

# stratification dates for each experiment
stratification <- read.table('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Setup/stratification_dates.txt', header = T)

# randomisation Files
randomFiles <- list.files('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Setup/GrowthChamber/', pattern = 'chamberRandomisation_.*_rep', full.names = T)
randomisations <- lapply(randomFiles, FUN = read.table, header = T, stringsAsFactors = F)
names(randomisations) <- c('16C_rep1', '16C_rep2', '16C_rep3', '6C_rep1', '6C_rep2', '6C_rep3')

# outliers
outliers <- read.table('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth/RawData/outlierList_allREps.txt', head = T)
```

```{r combine phenotype data}
# subset to relevant columns
overlapColnames <- c('Analysis.Label', 'Snapshot.ID.Tag', 'Snapshot.Time.Stamp', 'Area')
rep1.slim <- rep1[, c(overlapColnames)]
rep2_3.slim <- rep2_3[, c(overlapColnames)]

#combine
lemna <- rbind(rep1.slim, rep2_3.slim)
```

```{r experiment dates}
stratification[, (3:37)] <- apply(stratification[, c(3:37)], MARGIN = 2, function(x){as.POSIXlt(as.character(x), format = '%d.%m.%y')})
rep1_6C_start <- stratification$DAS14[stratification$rep == 'rep1' & stratification$temp == '6C']
rep1_6C_end <- stratification$DAS35[stratification$rep == 'rep1' & stratification$temp == '6C']
rep1_16C_start <- stratification$DAS14[stratification$rep == 'rep1' & stratification$temp == '16C']
rep1_16C_end <- stratification$DAS35[stratification$rep == 'rep1' & stratification$temp == '16C']

rep2_6C_start <- stratification$DAS14[stratification$rep == 'rep2' & stratification$temp == '6C']
rep2_6C_end <- stratification$DAS35[stratification$rep == 'rep2' & stratification$temp == '6C']
rep2_16C_start <- stratification$DAS14[stratification$rep == 'rep2' & stratification$temp == '16C']
rep2_16C_end <- stratification$DAS35[stratification$rep == 'rep2' & stratification$temp == '16C']

rep3_6C_start <- stratification$DAS14[stratification$rep == 'rep3' & stratification$temp == '6C']
rep3_6C_end <- stratification$DAS35[stratification$rep == 'rep3' & stratification$temp == '6C']
rep3_16C_start <- stratification$DAS14[stratification$rep == 'rep3' & stratification$temp == '16C']
rep3_16C_end <- stratification$DAS35[stratification$rep == 'rep3' & stratification$temp == '16C']


```


```{r annotate meta information}
# annotate date and hour
lemna$date <- unlist(lapply(lemna$Snapshot.Time.Stamp, getDate))
lemna$hour <- unlist(lapply(lemna$Snapshot.Time.Stamp, getHour))

# annotate experiment
lemna$experiment <- unlist(mapply(lemna$date, lemna$hour, FUN = getExperiment))

# remove datepoints without experiment (date from first 6C experiment)
lemna <- lemna[!is.na(lemna$experiment), ]

# annotate replicate and temperature
lemna$replicate <- unlist(lapply(lemna$experiment, function(exp){strsplit(exp, split = '_')[[1]][2]}))
lemna$temperature <- unlist(lapply(lemna$experiment, function(exp){strsplit(exp, split = '_')[[1]][1]}))

# annotate days after stratification (DAS)
lemna$DAS_decimal <- unlist(mapply(lemna$rep, lemna$temp, lemna$date, lemna$hour, FUN = getDAS.decimal))
lemna$DAS <- floor(lemna$DAS_decimal)

# combine hesoID annotation and create pot identifiers
heso <- rbind(heso.rep1, heso.rep2_3)
heso$pot <- paste(heso$tray, heso$coordinate, sep = '_')

# annotate potID, tray and coordinate
lemna$pot <- heso$pot[match(lemna$Snapshot.ID.Tag, heso$hesoID)]
lemna$tray <- unlist(lapply(lemna$pot, function(pot){strsplit(pot, split = '_')[[1]][1]}))
lemna$coordinate <- unlist(lapply(lemna$pot, function(pot){strsplit(pot, split = '_')[[1]][2]}))

# annotate accession
lemna$acn <- unlist(mapply(lemna$pot, lemna$experiment, FUN = getAccession))
```


### remove outliers
Outliers are defined as individual plants, or specific datapoints of individual plants that follow a growth trajectory of decreasing size.
Also specific timepoints for a given individual where size suddenly drops dramatically were considered as outliers.
Decision was based on plotting rosette area over time for each individual plants and manually going through those plots.
```{r remove outliers}
# define oulierIDs
lemna$outlierID <- paste(lemna$pot, lemna$temperature, lemna$replicate, lemna$DAS, sep = '_')
outliers$outlierID <- paste(outliers$pot, outliers$temp, outliers$rep, outliers$DAS, sep = '_')

# remove outliers
lemna.NO <- lemna[!lemna$outlierID %in% outliers$outlierID, ]
```



```{r write annotated phenotypingFile}
write.table(lemna, file = '/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth/rawdata_combined_annotation.txt', quote = F, row.names = F, sep = '\t')
write.table(lemna.NO, file = '/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth/rawdata_combined_annotation_NO.txt', quote = F, row.names = F, sep = '\t')
```




