---
title: "PLS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = F)
library(tidyverse)
library(caret)
library(olsrr)
knitr::opts_knit$set(root.dir = '/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/')
```
```{r Data}
# growth data
growth <- read_csv('Data/Growth/GWAS/growthPheno.csv')
# worldclim data
worldclim <- read_csv('/Volumes/nordborg/user/pieter.clauw/Documents/Source/Accessions/2029_modified_MN_SH_wc2.0_30s_bilinear.csv')
#clim <- read_csv('/Volumes/nordborg/user/pieter.clauw/Documents/Source/Accessions/araclim.csv')
meta <- read_csv('/Volumes/nordborg/user/pieter.clauw/Documents/Source/Accessions/A.thaliana_master_accession_list_011215.csv')
```
```{r general variables}
temperatures <- c('16C', '6C')
phenotypes <- c('LGR', 'rosette35DAS')

acns.SWE <- growth$accession[growth$accession %in% meta$id[meta$country == 'SWE']]
```
```{r tidying}
#clim %>%
#  mutate(autumnTemp = rowMeans(select(., tmean_10, tmean_11, tmean_12)))

growth.long <- growth %>%
  # remove accesisons 7471 from analysis. this accession lacks known coordinates
  filter(growth$accession != 7471) %>%
  select(everything(), rosette35DAS_response = rosette35DAS_perRed) %>%
  pivot_longer(
    cols = contains('_'),
    names_to = c(".value", "temperature"),
    names_sep = '_')

growthClim <- inner_join(growth.long, worldclim, by = 'accession') %>%
  select(accession, rosette35DAS, LGR, temperature, everything())
```


## multi-colinearity
It is very likely that there will be multi-colinearity in the explanatory climate variables.
Especially between the seasonal summaries and the corresponding months, but also between months and between different variables and the the bioclimatic summaries of those variables.

First we will assess how big the problem is.
Since we certainly expect colinearity between months and seasonal summaries, we will split up the analyses between monthly and seasonal data.

```{r multicolinearity}

```






```{r PLS}
for (pheno in phenotypes)
{
  nonPheno <- phenotypes[phenotypes != pheno]
  for (temp in c(temperatures, 'response'))
  {
    
    growthClim.temp.pheno <- growthClim %>%
      filter(temperature == temp) %>%
      select(-accession, -nonPheno, -temperature, -longitude, -latitude)
      
    # Compile cross-validation settings
    set.seed(100)
    myfolds <- createMultiFolds(as.vector(growthClim.temp.pheno[, pheno]), k = 5, times = 10)
    control <- trainControl("repeatedcv", index = myfolds, selectionFunction = "oneSE")

    # Train PLS model
    mod1 <- train(as.formula(paste(pheno,'~ .')), data = growthClim.temp.pheno,
      method = "pls",
      metric = "RMSE",
      tuneLength = 60,
      trcontrol = trainControl('cv', number = 10),
      preProc = c("zv","center","scale"))
  

    
          #trControl = control,
    # Check CV profile
    plot(mod1)
    
    pdf(paste('Results/Growth/Environment/PLS_', pheno, '_', temp, '.pdf', sep = ''))
    print(plot(varImp(mod1), 10, main = "PLS-DA"))
    dev.off()
    
  }
  
}


```

