---
title: "growthParameters"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = F)
library(tidyverse)
library(nlme)
source('Rfunctions.R')
source('../Functions/H.r')
knitr::opts_knit$set(root.dir = "/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/")
```

```{r data}
# estimated marginal means fron mixed nmodel
growth <- read_delim('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/lemna_DAS_allReps_NO_logRosette_emmeans_V3_CorCAR1_correct.txt', delim = '\t')
lemna <- read_delim('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth/rawdata_combined_annotation_NO.txt', delim = '\t')

# models
baseModel <- get(load("/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/poly3_logRosette_NO_allReps_V3_corCAR1_correct.rda"))
expFixed <- get(load("/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/poly3_logRosette_NO_allReps_V3_corCAR1_expFixed.rda"))

# randomisation Files
randomFiles <- list.files('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Setup/GrowthChamber/', pattern = 'chamberRandomisation_.*_rep', full.names = T)
randomisations <- lapply(randomFiles, FUN = read.table, header = T, stringsAsFactors = F)
names(randomisations) <- c('16C_rep1', '16C_rep2', '16C_rep3', '6C_rep1', '6C_rep2', '6C_rep3')
```

```{r remove accessions}
# 8343 shows a bad model fit to the data
growth <- growth[growth$acn != 8343, ]
```

```{r general variables}
accessions <- unique(growth$acn)
temperatures <- unique(growth$temp)
DASs <- unique(lemna$DAS)
experiments <- unique(lemna$experiment)
replicates <- unique(lemna$replicate)
replicates.h <- c(replicates, 'overall')

```

```{r visuals}
tempCol <- c('blue', 'red')
repPch <- c(15:18)
repType <- c('b', 'b', 'b', 'l')
visTempRep <- data.frame('temperature' = rep(temperatures, each = length(replicates.h)), 'replicate' = rep(replicates.h, length(temperatures)), 'tempCol' = rep(tempCol, each = length(replicates.h)), 'repPch' = rep(repPch, length(temperatures)), repType = rep(repType, length(temperatures)))
```

```{r model predictions}
# prepare newdata for prediction
predictDataFrames <- lapply(experiments, FUN = predictData.PerExperiment)
predictData <- do.call(rbind, predictDataFrames)
predictData$acn <- as.factor(as.character(predictData$acn))

# predict
predictData$logArea <- predict(baseModel, newdata = predictData)

# add meta info
predictData$replicate <- unlist(lapply(predictData$experiment, function(exp){strsplit(exp, split = '_')[[1]][2]}))
```

# Growth Parameters
After phenotyping and normalising the data, we obtained rosette area estimates for each accession summarised for each experiment but also summarised for each temperature (over the three replicate experiments).

Now we want to find the best summary for growth.

## rosette areas at 35DAS have relatively high heritabilities (especially within specific experiments)
```{r calculate heritabilities}
# dataframes to store heritbailities based on raw data and predicted data respectively
h <- data.frame('DAS' = rep(DASs, length(temperatures) * length(replicates.h)), 'temperature' =  rep(temperatures, each = length(DASs) * length(replicates.h)), 'replicate' = rep(rep(replicates.h, each = length(DASs)), length(temperatures)), 'H' = NA, stringsAsFactors = F)


for (rep in replicates.h)
{  
  # select specific replicate data if necessary
  predictData.r <- predictData
  if (rep != 'overall')
  {
    predictData.r <- predictData[predictData$replicate == rep, ]
  }
  
  for (t in temperatures)
  {
    for (DAS in DASs)
    {
      predictData.r.t.d <- predictData.r[predictData.r$temperature == t & predictData.r$DAS == DAS, c('acn', 'logArea')]
      if (nrow(predictData.r.t.d) == 0){h$H[h$DAS == DAS & h$temperature == t & h$replicate == rep] <- NA}
      else {h$H[h$DAS == DAS & h$temperature == t & h$replicate == rep] <- H(predictData.r.t.d, print = F)[4]}
    }
  }
}
```
```{r plot heritabilities}
minH <- min(h$H, na.rm = T)
maxH <- max(h$H, na.rm = T)
par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
plot(NA,NA, xlim = c(DASs[1], DASs[length(DASs)]), ylim = c(minH, maxH), xlab = 'days after stratification (DAS)', ylab = 'broad sense heritability')
for (t in temperatures)
{
  for (rep in replicates.h)
  {
    # select H for given temperature and replicate
    h.t.r <- h[h$temperature == t & h$replicate == rep, ]
    # plot properties
    col <- visTempRep$tempCol[visTempRep$temperature == t]
    pch <- visTempRep$repPch[visTempRep$replicate == rep]
    type <- unique(visTempRep$repType[visTempRep$replicate == rep])
    #plot lines
    lines(h.t.r$DAS, h.t.r$H, col = col, pch = pch, type = type)
    # legend
    legend('topright', inset=c(-0.25,0), legend = c(temperatures, replicates.h), col = c(tempCol, rep('black', length(replicates.h))), pch = c(NA,NA, repPch[1:3], NA), lty = c(1,1, NA, NA, NA, 1))
  }
}
```



## which fraction of heritability is due to initial growth (14DAS)?


## which summary for growth over time has sufficiently high heritabilities
```{r data preparation}
predictData.wide <- as_tibble(predictData) %>%
  pivot_wider(values_from = logArea, names_from = DAS, names_prefix = 'logArea_DAS_')
```

### percent increase
```{r percent increase}
predictData.wide <- predictData.wide %>%
  mutate(percentIncrease = ((logArea_DAS_35 - logArea_DAS_14)/logArea_DAS_14) * 100)

plot(predictData.wide$logArea_DAS_35, predictData.wide$percentIncrease)
```



### linear growth
#### not corrected for initial growth
##### between first and last time points
```{r linear growth noncorrected 14_35}
getSlope.ind.noncorr.14.35 <- function(ID)
{
  getSlope.ind(ID, DASs = c(14, 35), growth = predictData)
}

predictData.wide$LGR_nonCorr_14_35 <- mapply(predictData.wide$ID, FUN = getSlope.ind.noncorr.14.35)
```

##### over all time points
```{r linear growth noncorrected all timepoints}
getSlope.ind.noncorr.all <- function(ID)
{
  getSlope.ind(ID, DASs = c(14:35), growth = predictData)
}

predictData.wide$LGR_nonCorr_all <- mapply(predictData.wide$ID, FUN = getSlope.ind.noncorr.all)
```

#### corrected for initial growth
##### between first and last time points
```{r linear growth corrected 14_35}
# TEST #
getSlope.ind.corr <- function(ID, DASs, growth)
{
  growth.ind <- growth[growth$ID == ID & growth$DAS %in% DASs, ]
  growth.ind$DAS14 <- growth$logArea[growth$ID == ID & growth$DAS == 14]
  if (any(is.na(growth.ind$logArea))){return(NA)}
  else{
    fit <- lm(logArea ~ DAS14 + DAS, data = growth.ind)
    return(fit$effects[2])
  }
}


```


##### over all time points



## Try growthrates package
```{r growthrates TEST}
# TODO:tidyverse
predictData.ind <- predictData[predictData$ID == predictData$ID[1],]
predictData.ind$area <- exp(predictData.ind$logArea)
predictData.ind$DAS0 <- predictData.ind$DAS - 15

p     <- c(y0 = min(predictData.ind$area), mumax = 0.5)
lower <- c(y0 = min(predictData.ind$area) - 500, mumax = 0)
upper <- c(y0 = min(predictData.ind$area) + 500, mumax = 1)

fit_exp <- fit_growthmodel(FUN = grow_exponential, p = p, predictData.ind$DAS0, predictData.ind$area,
                        lower = lower, upper = upper)


p     <- c(y0 = min(predictData.ind$area), mumax = 0.5, K = max(predictData.ind$area)+20000)
lower <- c(y0 = min(predictData.ind$area) - 500, mumax = 0, K = max(predictData.ind$area) - 10000)
upper <- c(y0 = min(predictData.ind$area) + 500, mumax = 1, K = max(predictData.ind$area) + 30000)

fit_log <- fit_growthmodel(FUN = grow_logistic, p = p, predictData.ind$DAS0, predictData.ind$area,
                        lower = lower, upper = upper)


p     <- c(yi = 0.02, ya = 0.001, kw = 0.1, mumax = 0.2, K = 0.1)
lower <- c(yi = 1e-6, ya = 1e-6, kw = 0,    mumax = 0,   K = 0)
upper <- c(yi = 0.05, ya = 0.05, kw = 10,   mumax = 5,   K = 0.5)

fit2 <- fit_growthmodel(FUN = grow_twostep, p = p, time = dat$time, y = dat$value,
                        lower = lower, upper = upper)

coef(fit1)


```



## correlations between growth parameters


## heritabilities of growth parameters
