---
title: "indvidual nonlinear models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = F)
library(tidyverse)
library(nlme)
library(ggpubr)
knitr::opts_knit$set(root.dir = "/groups/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/")
```

## Modelling by individual
In order to calculate heritbaility estimates, a nonlinear model is built for each individual plant

## DISCLAIMER
Modelling per individual is giving NA vlaues for many individuals.
Better to include random effects fro each individual for M0, r and beta.
This is there for 16C in the fit2f.nlme.16C model.
Not yet for 6C. try to get these as well.

```{r data}
lemna <- read_delim('Data/Growth/rawdata_combined_annotation_NO.txt', delim = '\t')
load('Results/Growth/nonlinear/fit5.nlme.6C.rda')
load('Results/Growth/nonlinear/fit2f.nlme.16C.rda')

fit.nlme <- list('6C' = fit5.nlme.6C, '16C' = fit2f.nlme.16C)
```

```{r functions}
## self start function from Paine et al., 2012. Methods Ecol. Evol.
# Self-start function for power-law fit. This self-start for this model is not provided by Pinhero and Bates (2000), so I wrote one. Unfortunately, it often does not lead to convergence. Any suggestions would be welcome
fmla.pow <- as.formula("~(M0^(1-beta) + r*x*(1-beta))^(1/(1-beta))")
Init.pow <- function(mCall, LHS, data){
    xy <- sortedXyData(mCall[["x"]], LHS, data)
    if(nrow(xy) < 4) {stop("Too few distinct x values to fit a power-law")}
	r    <-coef(lm(log(y) ~ x, xy))[2]    # Use the slope from a log fit to the data as an initial guess for r
	M0   <- min(xy$y)		              # Use the minimum y value as an initial guess for M0
	beta <- 0.9	                          # give initial guess of beta as 0.9. don't use beta = 1, as it's undefined (1/0)
    value <- c(M0, r, beta) 
    names(value) <- mCall[c("M0", "r", "beta")]
 	return(value)
	}
SS.pow  <- selfStart(fmla.pow, initial = Init.pow, parameters = c("M0", "r", "beta"))

```

```{r data preparation}
lemna$ID <- paste(lemna$pot, lemna$experiment, sep = '_')
lemna$acn <- as.character(lemna$acn)
lemna$Area[lemna$Area == 0] <- NA

# only use accessions which have data for all replicates
acns_rep1 <- unique(lemna$acn[lemna$replicate == 'rep1'])
acns_rep2 <- unique(lemna$acn[lemna$replicate == 'rep2'])
acns_rep3 <- unique(lemna$acn[lemna$replicate == 'rep3'])

acns_rep123 <- intersect(intersect(acns_rep2, acns_rep3), acns_rep1)


lemna <- lemna %>%
  drop_na(Area) %>%
  filter(acn %in% acns_rep123) %>%
  select(ID, acn, experiment, temperature, replicate, DAS_decimal, Area) %>%
  mutate(DAS = round(DAS_decimal),
         ID = as.factor(ID),
         acn = as.factor(acn),
         temperature = as.factor(temperature),
         replicate = as.factor(replicate),
         experiment = as.factor(experiment))
```


```{r model per individual - nlsList}
ctrl <- nls.control(maxiter = 1000, warnOnly = T)
fm.lis <- nlsList(Area ~ SS.pow(DAS_decimal, M0, r, beta) | ID,
                         data = lemna,
                         control = ctrl)
```
produces many NA values

try with for loop
```{r model per individual}
ctrl <- nls.control(maxiter = 1000, warnOnly = T)
A.1_a1_16C_rep3           NA            NA            NA
A.1_a1_6C_rep1  1.284910e+03  2.330741e-02  1.116723e+00

ID <- 'A.1_a1_16C_rep3'
temp <- strsplit(ID, split = '_')[[1]][3]
acn <- unique(lemna$acn[lemna$ID == ID])

fix.temp <- fixef(fit.nlme[[temp]])
M0.start <- fix.temp['M0.(Intercept)'] + fix.temp[paste('M0.acn', acn, sep = '')]
r.start <- fix.temp['r.(Intercept)'] + fix.temp[paste('r.acn', acn, sep = '')]
beta.start <- fix.temp['beta']

lemna.ID <- filter(lemna, ID == !!ID)  
fit.ID <- nls(Area ~ SS.pow(DAS_decimal, M0, r, beta),
              start = list('M0' = M0.start, 'r' = r.start, 'beta' = beta.start),
              data = lemna.ID,
              control = ctrl,
              trace = T)


```




