---
title: "nonlinear_perID"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = F)
library(tidyverse)
library(furrr)
library(nlme)
library(ggpubr)
library(wesanderson)
#library(brms)
library(emmeans)
library(gridExtra)
library(bayesplot)
knitr::opts_knit$set(root.dir = "/groups/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/")
```

```{r data}
lemna <- read_delim('Data/Growth/rawdata_combined_annotation_NO.txt', delim = '\t')
outliers <- read_delim('Data/Growth/RawData/outlierList_nonlinear_perID.txt', delim = '\t')
# number of accessions
#n <- 20
```

```{r preparation}
# data preparation #
lemna$ID <- paste(lemna$pot, lemna$experiment, sep = '_')
lemna$acn <- as.character(lemna$acn)
lemna$Area[lemna$Area == 0] <- NA

# only use accessions which have data for all replicates
acns_rep1 <- unique(lemna$acn[lemna$replicate == 'rep1'])
acns_rep2 <- unique(lemna$acn[lemna$replicate == 'rep2'])
acns_rep3 <- unique(lemna$acn[lemna$replicate == 'rep3'])

acns_rep123 <- intersect(intersect(acns_rep2, acns_rep3), acns_rep1)

make_outlierIDs <- function(ID, DASlst)
{
  DASs <- c(0:21)
  if(DASlst != 'all'){DASs <- strsplit(DASlst, split = ';')[[1]]}
  outlier.ID <- paste(ID, DASs, sep = '_')
  return (outlier.ID)
}

# prepare outlier data for filtering
outlierIDs <- outliers %>%
  filter(remove == 'yes') %>%
  map2(.x = .$ID, .y = .$days, .f = ~make_outlierIDs(.x, .y)) %>%
  unlist()


# subset #
lemna <- lemna %>%
  filter(acn %in% acns_rep123) %>%
  drop_na(Area) %>%
  mutate(
         ID,
         acn = as.factor(acn),
         experiment = as.factor(experiment),
         temperature = as.factor(temperature),
         replicate = as.factor(replicate),
         DAS_decimal = DAS_decimal - 14,
         DAS = floor(DAS_decimal),
         Area,
         outlierID = paste(ID, DAS, sep = '_')) %>%
  filter(!outlierID %in% outlierIDs) %>%
  dplyr::select(ID, acn, experiment, temperature, replicate, DAS_decimal, DAS, Area)

# Area transformations
mean.t0 <- mean(lemna$Area[lemna$DAS == 0])
sd.t0 <- sd(lemna$Area[lemna$DAS == 0])
lemna <- lemna %>%
  mutate(Area.2000px = (Area/2000))

# summarise metadata for each ID
ID.meta <- lemna %>%
  dplyr::select(ID, acn, experiment, temperature, replicate) %>%
  distinct()
```

```{r general variables}
temperatures <- unique(lemna$temperature)
```


## per individual model fitting
### nlme
```{r SS.pow}
## self start function from Paine et al., 2012. Methods Ecol. Evol.
# Self-start function for power-law fit. This self-start for this model is not provided by Pinhero and Bates (2000), so I wrote one. Unfortunately, it often does not lead to convergence. Any suggestions would be welcome
fmla.pow <- as.formula("~(M0^(1-beta) + r*x*(1-beta))^(1/(1-beta))")
Init.pow <- function(mCall, LHS, data){
    xy <- sortedXyData(mCall[["x"]], LHS, data)
    if(nrow(xy) < 4) {stop("Too few distinct x values to fit a power-law")}
	r    <-coef(lm(log(y) ~ x, xy))[2]    # Use the slope from a log fit to the data as an initial guess for r
	M0   <- min(xy$y)		              # Use the minimum y value as an initial guess for M0
	beta <- 0.9	                          # give initial guess of beta as 0.9. don't use beta = 1, as it's undefined (1/0)
    value <- c(M0, r, beta) 
    names(value) <- mCall[c("M0", "r", "beta")]
 	return(value)
	}
SS.pow  <- selfStart(fmla.pow, initial = Init.pow, parameters = c("M0", "r", "beta"))
```


```{r nlsList}
ctrl.nls <- nls.control(maxiter = 1000, warnOnly = T)

fit.lis <- nlsList(Area.2000px ~ SS.pow(DAS_decimal, M0, r, beta) | ID,
                   data = lemna,
                   control = ctrl.nls.2)

fit.lis.coef <- as_tibble(coef(fit.lis), rownames = 'ID') %>%
  left_join(ID.meta, by = 'ID') %>%
  dplyr::select(ID, acn, experiment, temperature, replicate, M0, r, beta)

save(fit.lis, file = '/scratch-cbe/users/pieter.clauw/16vs6/Results/Growth/nonlinear/perID/models/nlme/fit.lis.rda')
save(fit.lis.coef, file = '/scratch-cbe/users/pieter.clauw/16vs6/Results/Growth/nonlinear/perID/models/nlme/fit.lis.coef.rda')
```

```{r results nlsList}
M0.hist <- fit.lis.coef %>% 
  ggplot(aes(x = M0, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.05) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

r.hist <- fit.lis.coef %>% 
  ggplot(aes(x = r, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.005) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

beta.hist <- fit.lis.coef %>% 
  ggplot(aes(x = beta, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.05) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

ggarrange(M0.hist, r.hist, beta.hist)
```

There are some plants with NA or outlier values.
check how many, and if this causes problems for certain accessions, replicates, temepratures.
Need for removal?

```{r gather NAs + outliers}
noFit.lis <- fit.lis.coef %>%
  filter(is.na(M0) | is.na(r) | is.na(beta))

table(noFit.lis$temperature, noFit.lis$replicate)

r.outlier.lis <- fit.lis.coef %>%
  filter(r > 2 | r < -2)

beta.outlier.lis <- fit.lis.coef %>%
  filter(beta < -2 | beta > 2)

misfits.lis <- unique(c(noFit.lis$ID, r.outlier.lis$ID, beta.outlier.lis$ID))
```

```{r plot NAs and outliers - misfits}
# plot data of NA and outlier IDs

for (ID in misfits.lis){

  ID.pred <- 0
  if(!is.na(fit.lis.coef$M0[fit.lis.coef$ID == ID])){
    ID.pred <- predict(fit.lis[[ID]])
  }
  
  lemna.ID <- filter(lemna, ID == !!ID) %>%
    arrange(DAS_decimal) %>%
    mutate(modelFit = ID.pred)
    
  p <- ggplot(lemna.ID, aes(DAS_decimal, Area.2000px)) +
    geom_point() +
    geom_line(aes(DAS_decimal, modelFit), colour = wes_palette('Darjeeling1', n = 1)) +
    ggtitle(ID) +
    theme_minimal()
  
  pdf(paste('/scratch-cbe/users/pieter.clauw/16vs6/Results/Growth/nonlinear/perID/misfits/misfit.data_', ID, '.pdf', sep = ''))
  print(p)
  dev.off()
  rm(lemna.ID)
}
```
The selected misfits above have different reasons for showing outlier or NA fits.
Among the NA fits we find IDs for which the data is clearly not following a power law.
For some plants there is a visible logistic growth pattern, other have weird patterns (bad phenotyping).

Among the plants with outlier fits, there are some which do look decent. Outlier criterium therefore must be relaxed.
Others show assymptotic growth adn some extreme cases show fractionated growth patterns.

Because there are priors about the dsitributions of M0, r and beta.
We also tried a bayesian approach and see if this creates less outlier model fits.
this however did not deliver results as expected (unrealsitic distribution of beta values), moreover, the estimated sampple size (ESS) was in many cases not satisfactory.
Therefore we continue with the nlsList results


Get the estimates per accession for each temperature
correct for replicate effects and/or experiment effects
especially M0 may have a stro,ng replicate effect (rep1 vs rep2&3)


```{r significance of experiment effects}
fit.lis.coef.NAfree <- na.omit(fit.lis.coef)
M0.lm <- lm(M0 ~ acn * temperature + experiment, data = fit.lis.coef.NAfree)
r.lm <- lm(r ~ acn * temperature + experiment, data = fit.lis.coef.NAfree)
beta.lm <- lm(beta ~ acn * temperature + experiment, data = fit.lis.coef.NAfree)

anova(M0.lm)
anova(r.lm)
anova(beta.lm)

M0.lm <- lm(M0 ~ acn * temperature * experiment, data = fit.lis.coef.NAfree)
r.lm <- lm(r ~ acn * temperature * experiment, data = fit.lis.coef.NAfree)
beta.lm <- lm(beta ~ acn * temperature * experiment, data = fit.lis.coef.NAfree)

anova(M0.lm)
anova(r.lm)
anova(beta.lm)

```


```{r get accession estimates}
fit.lis.coef.NAfree <- na.omit(fit.lis.coef)

# M0
M0.lme <- lme(M0 ~ acn *temperature, random = ~1 |experiment, data = fit.lis.coef.NAfree)
anova(M0.lme)
M0.noTemp.lme <- lme(M0 ~ acn , random = ~1 |experiment/ID, data = fit.lis.coef.NAfree)

M0.emm <- as_tibble(emmeans(M0.lme, ~ acn * temperature))
M0.noTemp.emm <- as_tibble(emmeans(M0.noTemp.lme, ~ acn)) %>%
  mutate(temperature = '16C') %>%
  bind_rows(., mutate(., temperature = '6C')) %>%
  select(acn, temperature, emmean, SE, df, lower.CL, upper.CL) %>%
  mutate(temperature = as.factor(temperature))

# r
r.lme <- lme(r ~ acn * temperature, random = ~1 | experiment, data= fit.lis.coef.NAfree)
anova(r.lme)

r.emm <- as_tibble(emmeans(r.lme, ~ acn * temperature))

# beta
beta.lme <- lme(beta ~ acn * temperature, random = ~1 | experiment, data= fit.lis.coef.NAfree)
anova(beta.lme)

beta.emm <- as_tibble(emmeans(beta.lme, ~ acn * temperature))

growth <- full_join(M0.emm, M0.noTemp.emm, by = c('acn', 'temperature'), suffix = c('.M0', '.M0.noTemp')) %>%
  full_join(., r.emm, by = c('acn', 'temperature')) %>%
  rename(r = emmean, SE.r = SE, df.r = df, lower.CL.r = lower.CL, upper.CL.r = upper.CL) %>%
  full_join(., beta.emm, by = c('acn', 'temperature')) %>%
  rename(accession = acn, M0 = emmean.M0, M0.noTemp = emmean.M0.noTemp, beta = emmean, SE.beta = SE, df.beta = df, lower.CL.beta = lower.CL, upper.CL.beta = upper.CL)

write_csv(growth, path = '/groups/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/nonlinear/growthParameters.csv')
```

```{r growth estimate distributions}
M0.hist <- growth %>% 
  ggplot(aes(x = M0, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.025) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

M0.noTemp.hist <- growth %>% 
  ggplot(aes(x = M0.noTemp, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.025) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

r.hist <- growth %>% 
  ggplot(aes(x = r, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.0025) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

beta.hist <- growth %>% 
  ggplot(aes(x = beta, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.025) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

ggarrange(M0.hist, M0.noTemp.hist, r.hist, beta.hist)
```


Plot the model for each accession and each temperature against the data.
(check ggplot for models)
```{r}
growth %>%
pmap(function(...){
  growth.par <- tibble(...)
  acn <- growth.par$accession
  temp <- growth.par$temperature
  M0 <- growth.par$M0
  M0.noTemp <- growth.par$M0.noTemp
  r <- growth.par$r
  beta <- growth.par$beta
  lemna.a.t <- filter(lemna, acn == !!acn, temperature == !!temp)
  pred.a.t <- tibble(DAS_decimal = seq(0.5, 21.5, length.out = 100)) %>%
    mutate(Area.2000px.pred = (M0^(1-beta) + r*DAS_decimal*(1-beta))^(1/(1-beta)),
           Area.2000px.M0.noTemp.pred = (M0.noTemp^(1-beta) + r*DAS_decimal*(1-beta))^(1/(1-beta)))
  
  p <- ggplot(aes(DAS_decimal, Area.2000px), data = lemna.a.t) +
    geom_point(aes(col = replicate)) +
    geom_line(aes(DAS_decimal, Area.2000px.pred), data = pred.a.t) +
    scale_color_manual(values = wes_palette('Darjeeling1')) +
    theme_minimal() +
    labs(title = paste(acn, temp), x = 'days', y = 'rosette area')
  
  p.M0.noTemp <- ggplot(aes(DAS_decimal, Area.2000px), data = lemna.a.t) +
    geom_point(aes(col = replicate)) +
    geom_line(aes(DAS_decimal, Area.2000px.M0.noTemp.pred), data = pred.a.t) +
    scale_color_manual(values = wes_palette('Darjeeling1')) +
    theme_minimal() +
    labs(title = paste(acn, temp), x = 'days', y = 'rosette area')
  
  pdf(paste('/scratch-cbe/users/pieter.clauw/16vs6/Results/Growth/nonlinear/perID/modelFits_acn_temp/fit_', acn, '_', temp, '.pdf', sep = ''))
  print(p)
  dev.off()
  
    pdf(paste('/scratch-cbe/users/pieter.clauw/16vs6/Results/Growth/nonlinear/perID/modelFits_acn_temp/fit_M0.noTemp_', acn, '_', temp, '.pdf', sep = ''))
  print(p.M0.noTemp)
  dev.off()
})
```


```{r temperature responses}
growth.temp <- growth %>%
  pivot_wider()
```




Calculate heritabilities
```{r}

```







## Try bayesian analysis with reasonably flexible priors
Bayesian per individual plant analyses were run using nonlinear_bayesian_perID_cluster2.0.sh
```{r summarise bayesian results}
fit.bayes.coef <- tibble(
  ID = character(),
  M0 = numeric(),
  r = numeric(),
  beta = numeric(),
  M0_ESS = numeric(),
  r_ESS = numeric(),
  beta_ESS = numeric())
missingAcns <- vector()
for (acn in acns_rep123)
{
  file.acn <- paste('/groups/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/nonlinear/models/coef_perAcn_2.0/fit.bayesian_2.0.coef_', acn, '.csv', sep = '')
  if (! file.exists(file.acn)){warning(paste('no summary file for accession:', acn)); missingAcns <- c(missingAcns, acn); next}
  coef.acn <- read_csv(file.acn)
  fit.bayes.coef <- bind_rows(fit.bayes.coef, coef.acn)
}

fit.bayes.coef <- fit.bayes.coef %>%
  left_join(ID.meta, by = 'ID') %>%
  dplyr::select(ID, acn, experiment, temperature, replicate, M0, r, beta, M0_ESS, r_ESS, beta_ESS)

```

```{r results bayesian}
M0.hist <- fit.bayes.coef %>% 
  ggplot(aes(x = M0, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.05) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

r.hist <- fit.bayes.coef %>% 
  ggplot(aes(x = r, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.005) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

beta.hist <- fit.bayes.coef %>% 
  ggplot(aes(x = beta, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.005) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

ggarrange(M0.hist, r.hist, beta.hist)
```
```{r ESS}
M0_ESS.hist <- fit.bayes.coef %>% 
  ggplot(aes(x = M0_ESS, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity') +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

r_ESS.hist <- fit.bayes.coef %>% 
  ggplot(aes(x = r_ESS, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity') +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

beta_ESS.hist <- fit.bayes.coef %>% 
  ggplot(aes(x = beta_ESS, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity') +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

ggarrange(M0_ESS.hist, r_ESS.hist, beta_ESS.hist)
```
```{r ESS correlations}
M0_r_ESS <- fit.bayes.coef %>%
  ggplot(aes(x = M0_ESS, y = r_ESS, color = temperature)) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

M0_beta_ESS <- fit.bayes.coef %>%
  ggplot(aes(x = M0_ESS, y = beta_ESS, color = temperature)) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

r_beta_ESS <- fit.bayes.coef %>%
  ggplot(aes(x = r_ESS, y = beta_ESS, color = temperature)) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

ggarrange(M0_r_ESS, M0_beta_ESS, r_beta_ESS)
```
```{r gather NAs + bad estimates}
noFit.bayes <- fit.bayes.coef %>%
  filter(is.na(M0) | is.na(r) | is.na(beta))

table(noFit.bayes$temperature, noFit.bayes$replicate)


r.outlier.bayes <- fit.bayes.coef %>%
  filter(r > 2 | r < -2)

beta.outlier.bayes <- fit.bayes.coef %>%
  filter(beta < -2 | beta > 2)

misfits.bayes <- unique(c(noFit.bayes$ID, r.outlier.bayes$ID, beta.outlier.bayes$ID))
```

```{r significance of experiment effects}
fit.bayes.coef.NAfree <- na.omit(fit.bayes.coef)
M0.lm <- lm(M0 ~ acn * temperature + experiment, data = fit.bayes.coef.NAfree)
r.lm <- lm(r ~ acn * temperature + experiment, data = fit.bayes.coef.NAfree)
beta.lm <- lm(beta ~ acn * temperature + experiment, data = fit.bayes.coef.NAfree)

anova(M0.lm)
anova(r.lm)
anova(beta.lm)

M0.glm <- glm(M0 ~ acn * temperature + experiment, data = fit.bayes.coef.NAfree, family = gaussian)
r.lm <- lm(r ~ acn * temperature + experiment, data = fit.bayes.coef.NAfree)
beta.lm <- lm(beta ~ acn * temperature + experiment, data = fit.bayes.coef.NAfree)


M0.lm <- lm(M0 ~ acn * temperature * experiment, data = fit.bayes.coef.NAfree)
r.lm <- lm(r ~ acn * temperature * experiment, data = fit.bayes.coef.NAfree)
beta.lm <- lm(beta ~ acn * temperature * experiment, data = fit.bayes.coef.NAfree)

anova(M0.lm)
anova(r.lm)
anova(beta.lm)
```

```{r temperature effects}
fit.bayes.coef.NAfree <- na.omit(fit.bayes.coef)
M0.lme <- lme(M0 ~ acn + temperature, random = ~1|experiment, data = fit.bayes.coef.NAfree)
r.lme <- lme(r ~ acn + temperature, random = ~1|experiment, data = fit.bayes.coef.NAfree)
beta.lme <- lme(beta ~ acn + temperature, random = ~1|experiment, data = fit.bayes.coef.NAfree)

M0.aov <- anova(M0.lme)
r.aov <- anova(r.lme)
beta.aov <- anova(beta.lme)

sign.levels <- c(-1, 0.001, 0.01, 0.05, 1)
sign.labels <- c('<0.001', '< 0.01','<0.05','NS')

tble.df <- tibble(terms = rownames(M0.aov), M0 = cut(M0.aov$`p-value`, sign.levels, sign.labels), r = cut(r.aov$`p-value`, sign.levels, sign.labels), beta = cut(beta.aov$`p-value`, sign.levels, sign.labels))
  
pdf('/groups/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/nonlinear/temperatureEffect_anova_2.0.pdf')
grid.table(tble.df[, -1], rows = tble.df$terms)
dev.off()
```
```{r accession x temperature effect}
fit.bayes.coef.NAfree <- na.omit(fit.bayes.coef)
M0.lme <- lme(M0 ~ acn * temperature, random = ~1|experiment, data = fit.bayes.coef.NAfree)
r.lme <- lme(r ~ acn * temperature, random = ~1|experiment, data = fit.bayes.coef.NAfree)
beta.lme <- lme(beta ~ acn * temperature, random = ~1|experiment, data = fit.bayes.coef.NAfree)

M0.aov <- anova(M0.lme)
r.aov <- anova(r.lme)
beta.aov <- anova(beta.lme)

sign.levels <- c(-1, 0.001, 0.01, 0.05, 1)
sign.labels <- c('<0.001', '< 0.01','<0.05','NS')

tble.df <- tibble(terms = rownames(M0.aov), M0 = cut(M0.aov$`p-value`, sign.levels, sign.labels), r = cut(r.aov$`p-value`, sign.levels, sign.labels), beta = cut(beta.aov$`p-value`, sign.levels, sign.labels))

pdf('/groups/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/nonlinear/temperatureInteractionEffect_anova_2.0.pdf')
grid.table(tble.df[, -1], rows = tble.df$terms)
dev.off()

```

No temperature effects for M0 and beta.
Only in interaction with accession.


```{r estimates per acn}
fit.bayes.coef.NAfree <- na.omit(fit.bayes.coef)

# M0
M0.lme <- lme(M0 ~ acn * temperature, random = ~1|experiment, data = fit.bayes.coef.NAfree)
M0.noTemp.lme <- lme(M0 ~ acn , random = ~1|experiment, data = fit.bayes.coef.NAfree)

M0.emm <- as_tibble(emmeans(M0.lme, ~ acn * temperature))

M0.noTemp.emm <- as_tibble(emmeans(M0.noTemp.lme, ~ acn)) %>%
  mutate(temperature = '16C') %>%
  bind_rows(., mutate(., temperature = '6C')) %>%
  select(acn, temperature, emmean, SE, df, lower.CL, upper.CL) %>%
  mutate(temperature = as.factor(temperature))

# r
r.lme <- lme(r ~ acn * temperature, random = ~1|experiment, data= fit.bayes.coef.NAfree)

r.emm <- as_tibble(emmeans(r.lme, ~ acn * temperature))

# beta
beta.lme <- lme(beta ~ acn * temperature, random = ~1|experiment, data= fit.bayes.coef.NAfree)
beta.noTemp.lme <- lme(beta ~ acn , random = ~1|experiment, data = fit.bayes.coef.NAfree)

beta.emm <- as_tibble(emmeans(beta.lme, ~ acn * temperature))
beta.noTemp.emm <- as_tibble(emmeans(beta.noTemp.lme, ~ acn)) %>%
  mutate(temperature = '16C') %>%
  bind_rows(., mutate(., temperature = '6C')) %>%
  select(acn, temperature, emmean, SE, df, lower.CL, upper.CL) %>%
  mutate(temperature = as.factor(temperature))

growth <- full_join(M0.emm, M0.noTemp.emm, by = c('acn', 'temperature'), suffix = c('.M0', '.M0.noTemp')) %>%
  full_join(., r.emm, by = c('acn', 'temperature')) %>%
  rename(r = emmean, SE.r = SE, df.r = df, lower.CL.r = lower.CL, upper.CL.r = upper.CL) %>%
  full_join(., beta.emm, by = c('acn', 'temperature')) %>%
  rename(beta = emmean, SE.beta = SE, df.beta = df, lower.CL.beta = lower.CL, upper.CL.beta = upper.CL) %>%
  full_join(., beta.noTemp.emm, by = c('acn', 'temperature')) %>%
  rename(accession = acn, M0 = emmean.M0, M0.noTemp = emmean.M0.noTemp, beta.noTemp = emmean, SE.beta.noTmep = SE, df.beta.noTemp = df, lower.CL.beta.noTemp = lower.CL, upper.CL.beta.noTemp = upper.CL)

# add rosette area at final day.
growth <- growth %>%
  mutate(Area.DAS21 = (M0^(1-beta) + r*21*(1-beta))^(1/(1-beta)),
         RGR.DAS21 = r*(M0^(1-beta) + r*21*(1-beta))^-1)

write_csv(growth, path = '/groups/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/nonlinear/growthParameters_bayesian_2.0.csv')
```


```{r growth estimate distributions}
M0.hist <- growth %>% 
  ggplot(aes(x = M0, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.025) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

M0.noTemp.hist <- growth %>% 
  ggplot(aes(x = M0.noTemp, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.025) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

r.hist <- growth %>% 
  ggplot(aes(x = r, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.0025) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

beta.hist <- growth %>% 
  ggplot(aes(x = beta, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.00025) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

beta.noTemp.hist <- growth %>% 
  ggplot(aes(x = beta.noTemp, fill = temperature)) +
  geom_histogram(alpha = 0.6, position = 'identity', binwidth = 0.025) +
  scale_fill_manual(values = wes_palette('Darjeeling1')) +
  theme_minimal()

ggarrange(M0.hist, M0.noTemp.hist, r.hist, beta.hist)
pdf('/groups/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/nonlinear/growthParameters_M0_r_beta_2.0.pdf')
ggarrange(M0.hist, r.hist, beta.hist)
dev.off()
```

```{r plot estimated rosette areas per accession}
growth %>%
pmap(function(...){
  growth.par <- tibble(...)
  acn <- growth.par$accession
  temp <- growth.par$temperature
  M0 <- growth.par$M0
  M0.noTemp <- growth.par$M0.noTemp
  r <- growth.par$r
  beta <- growth.par$beta
  lemna.a.t <- filter(lemna, acn == !!acn, temperature == !!temp)
  pred.a.t <- tibble(DAS_decimal = seq(0.5, 21.5, length.out = 100)) %>%
    mutate(Area.2000px.pred = (M0^(1-beta) + r*DAS_decimal*(1-beta))^(1/(1-beta)),
           Area.2000px.M0.noTemp.pred = (M0.noTemp^(1-beta) + r*DAS_decimal*(1-beta))^(1/(1-beta)))
  
  p <- ggplot(aes(DAS_decimal, Area.2000px), data = lemna.a.t) +
    geom_point(aes(col = replicate)) +
    geom_line(aes(DAS_decimal, Area.2000px.pred), data = pred.a.t) +
    scale_color_manual(values = wes_palette('Darjeeling1')) +
    theme_minimal() +
    labs(title = paste(acn, temp), x = 'days', y = 'rosette area')
  
  p.M0.noTemp <- ggplot(aes(DAS_decimal, Area.2000px), data = lemna.a.t) +
    geom_point(aes(col = replicate)) +
    geom_line(aes(DAS_decimal, Area.2000px.M0.noTemp.pred), data = pred.a.t) +
    scale_color_manual(values = wes_palette('Darjeeling1')) +
    theme_minimal() +
    labs(title = paste(acn, temp), x = 'days', y = 'rosette area')
  
  pdf(paste('/scratch-cbe/users/pieter.clauw/16vs6/Results/Growth/nonlinear/perID/modelFits_acn_temp/bayesian2.0/fit_', acn, '_', temp, '.pdf', sep = ''))
  print(p)
  dev.off()
  
    #pdf(paste('/scratch-cbe/users/pieter.clauw/16vs6/Results/Growth/nonlinear/perID/modelFits_acn_temp/fit_M0.noTemp_', acn, '_', temp, '.pdf', sep = ''))
  #print(p.M0.noTemp)
  #dev.off()
})
```

```{r heritabilities}
phenotypes <- c('M0', 'r', 'beta')
growth.herit <- data.frame(phenotype = rep(phenotypes, length(temperatures)), temperature = rep(temperatures, each = length(phenotypes)), H = NA)

for (temp in temperatures)
{
  for (pheno in c('M0', 'r', 'beta'))
  {
    fit.coef <- fit.bayes.coef %>%
      select(acn, temperature, experiment, value = !!pheno) %>%
      filter(temperature == temp, !is.na(value))
    
    lmm <- lme(value ~ 1, random = ~ 1|experiment/acn, data = fit.coef)
    theta_lmm <- as.numeric(VarCorr(lmm)[4:5])
    H <- theta_lmm[1] / (theta_lmm[1] + theta_lmm[2])
    
    growth.herit$H[growth.herit$phenotype == pheno & growth.herit$temperature == temp] <- H
  }
}




herit.tble <- growth.herit %>%
  mutate(H = round(H, digits = 2))

pdf('/groups/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/nonlinear/broadSenseHeritabilities_2.0.pdf')
grid.table(herit.tble[, -1], rows = herit.tble$phenotype)
dev.off()
```

```{r growth phenotypes}
growth.pheno <- growth %>%
  select(accession, temperature, M0, M0.noTemp, r, beta, beta.noTemp, Area.DAS21, RGR.DAS21) %>%
  pivot_wider(id_cols = accession, names_from = temperature, values_from = c('M0', 'M0.noTemp', 'r', 'beta', 'beta.noTemp', 'Area.DAS21', 'RGR.DAS21'), names_sep = '_')

# response slopes
resp.curve <- function(pheno_6C, pheno_16C)
{
  pheno <- tibble('X_6C' = pheno_6C, 'X_16C' = pheno_16C) %>%
    pivot_longer(cols = starts_with('X'), names_to = 'temperature', names_prefix = 'X_', values_to = 'pheno') %>%
    mutate(temperature = factor(temperature, levels = c('6C', '16C'), ordered = T))
  fit <- lm(pheno ~ temperature, data = pheno)
  resp <- as.numeric(fit$coefficients[2])
  return(resp)
}

growth.pheno <- growth.pheno %>%
  mutate(M0_response = unlist(map2(.x = M0_6C, .y = M0_16C, ~ resp.curve(.x, .y))),
         r_response = unlist(map2(.x = r_6C, .y = r_16C, ~ resp.curve(.x, .y))),
         beta_response = unlist(map2(.x = beta_6C, .y = beta_16C, ~ resp.curve(.x, .y))),
         Area.DAS21_response = unlist(map2(.x = Area.DAS21_6C, .y = Area.DAS21_16C, ~ resp.curve(.x, .y))),
         RGR.DAS21_response = unlist(map2(.x = RGR.DAS21_6C, .y = RGR.DAS21_16C, ~ resp.curve(.x, .y)))) %>%
  mutate(M0.noTemp = M0.noTemp_16C, beta.noTemp = beta.noTemp_16C) %>%
  select(-M0.noTemp_16C, -beta.noTemp_16C, -M0.noTemp_6C, -beta.noTemp_6C)

write_csv(growth.pheno, path = '/groups/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/nonlinear/growthPhenotypes_bayesian_2.0.csv')
```


