---
title: "FixingHesoIDs"
author: "Pieter Clauw"
date: "11/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = F)
source('/Volumes/nordborg/user/pieter.clauw/Documents/Source/Rfunctions/H.r')
```

## The solution?
From the pictures it seems that the positions annotated to the hesoIDs are wrong.
In teh trays, the indication for filling the trays indicates that columns 'a' to 'e' are filled from left to right, while the annotation with the hesoIDs follows columns 'a' to 'e' from right to left.

```{r data}
heso.rep2_3 <- read.table('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth.old//RawData/hesoAnnotations_MartoJune_2018.txt', header = T, sep = '\t')
stratification <- read.table('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Setup/stratification_dates.txt', header = T)
rep2_3 <- read.table('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth.old/RawData/LemnaTecResults_16v62018_2017Analysis.csv', header = T, sep = ';')

randomFiles <- list.files('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Setup/GrowthChamber/', pattern = 'chamberRandomisation_.*_rep', full.names = T)
randomisations <- lapply(randomFiles, FUN = read.table, header = T, stringsAsFactors = F)
names(randomisations) <- c('16C_rep1', '16C_rep2', '16C_rep3', '6C_rep1', '6C_rep2', '6C_rep3')
```


```{r functions}
getDate <- function(timeStamp)
{
  if (grepl('[A,P]M', timeStamp))
  {return(format(strptime(timeStamp, format = '%m/%d/%Y %I:%M:%S %p'), '%d.%m.%Y'))}
  else if (!grepl('[A,P]M', timeStamp))
  {return(format(strptime(timeStamp, format = '%d.%m.%Y %H:%M:%S'), '%d.%m.%Y'))}
}

getHour <- function(timeStamp)
{
  if (grepl('[A,P]M', timeStamp))
  {return(format(strptime(timeStamp, format = '%m/%d/%Y %I:%M:%S %p'), '%H:%M'))}
  else if (!grepl('[A,P]M', timeStamp))
  {return(format(strptime(timeStamp, format = '%d.%m.%Y %H:%M:%S'), '%H:%M'))}
}

getExperiment <- function(date, hour)
{
  date <- strptime(date, format = '%d.%m.%Y')
  hour <- strptime(hour, format = '%H:%M')
  noon <- strptime('12:00', format = '%H:%M')
  if (date >= rep1_6C_start & date <= rep1_6C_end){
    return('6C_rep1')
  }  else if (date >= rep1_16C_start & date <= rep1_16C_end){
    return('16C_rep1')
  } else if (date >= rep2_6C_start & date < rep2_6C_end){
    return('6C_rep2')
  }  else if (date == rep2_6C_end & hour < noon){
    return('6C_rep2')
  }  else if (date == rep2_16C_start & hour > noon){
    return('16C_rep2')
  }  else if (date > rep2_16C_start & date < rep2_16C_end){
    return('16C_rep2')
  }  else if (date == rep2_16C_end & hour < noon){
    return('16C_rep2')
  }  else if (date == rep3_6C_start & hour > noon){
    return('6C_rep3')
  }  else if (date > rep3_6C_start & date < rep3_6C_end){
    return('6C_rep3')
  }  else if (date == rep3_6C_end & hour < noon){
    return('6C_rep3')
  }  else if (date == rep3_16C_start & hour > noon){
    return('16C_rep3')
  }  else if (date > rep3_16C_start & date <= rep3_16C_end){
    return('16C_rep3')
  } else {return(NA)}
}

getDAS.decimal <- function(rep, temp, date, hour)
{
  #if(rep == 'rep1_re'){rep <- 'rep1'}
  if (grepl('_re', rep)){rep <- strsplit(rep, split = '_')[[1]][1]}
  DAS1 <- stratification$DAS1[stratification$temp == temp & stratification$rep == rep]
  datehour <- paste(date, hour)
  datehour <- as.POSIXlt(datehour, format = '%d.%m.%Y %H:%M')
  DAS.dec <- as.numeric(difftime(datehour, DAS1, units = 'days')) + 1
  if (length(DAS.dec) == 0){DAS.dec <- NA}
  return(DAS.dec)
}

getAccession <- function(pot, exp)
{
  if (grepl('rep[1,2,3]_re', exp)){exp <- paste(strsplit(exp, split = '_')[[1]][c(1,2)], collapse = '_')}
  random.e <- randomisations[[exp]]
  random.e$pot <- paste(random.e$tray, random.e$coord, sep = '_')
  acn <- random.e$acn[random.e$pot == pot]  
  return(acn)
}

getH_perExp_perDAS <- function(exp, DAS)
{
  hData <- rep2_3.slim[rep2_3.slim$experiment == exp & rep2_3.slim$DAS == DAS, c('acn', 'Area')]
  if (all(is.na(hData$Area))) {return(NA)}
  else {return(H(hData)[4])}
}

getVe_perExp_perDAS <- function(exp, DAS)
{
  hData <- rep2_3.slim[rep2_3.slim$experiment == exp & rep2_3.slim$DAS == DAS, c('acn', 'Area')]
  if (all(is.na(hData$Area))) {return(NA)}
  else {return(Ve(hData))}
}

getVg_perExp_perDAS <- function(exp, DAS)
{
  hData <- rep2_3.slim[rep2_3.slim$experiment == exp & rep2_3.slim$DAS == DAS, c('acn', 'Area')]
  if (all(is.na(hData$Area))) {return(NA)}
  else {return(Vg(hData))}
}

```



## Fix-it
Adjust the hesoID annotation file, so that the corerct positions are linked to the hesoIDs.

```{r fix heso annotation}


heso.rep2_3.correct <- heso.rep2_3
hesoCoordKey <- data.frame('wrong' = c('e', 'd', 'c', 'b', 'a'), 'right' = c('a', 'b', 'c', 'd', 'e'))
for (i in 1:nrow(heso.rep2_3.correct))
{
  c.wrong <- heso.rep2_3$coordinate[i]
  c.wrong.col <- substr(c.wrong, 1,1)
  c.wrong.row <- substr(c.wrong, 2,2)
  c.right.col <- hesoCoordKey$right[hesoCoordKey$wrong == c.wrong.col]
  c.right <- paste(c.right.col, c.wrong.row, sep = '')
  
  heso.rep2_3.correct$coordinate[i] <- c.right
}

write.table(heso.rep2_3.correct, file = '/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Data/Growth/hesoAnnotations_MartoJune_2018_correct.txt', row.names = F, quote = F, sep = '\t')
```

## reannotate
time to make some pudding
```{r reannotate}
rep2_3.slim <- rep2_3[, c('Analysis.Label', 'Snapshot.ID.Tag', 'Snapshot.Time.Stamp', 'Area')]

# add date and hour
rep2_3.slim$date <- unlist(lapply(rep2_3.slim$Snapshot.Time.Stamp, getDate))
rep2_3.slim$hour <- unlist(lapply(rep2_3.slim$Snapshot.Time.Stamp, getHour))

# change stratification dates iinto POSIXlt format
stratification[, (3:37)] <- apply(stratification[, c(3:37)], MARGIN = 2, function(x){as.POSIXlt(as.character(x), format = '%d.%m.%y')})

# start and end dates for each experiment
rep1_6C_start <- stratification$DAS14[stratification$rep == 'rep1' & stratification$temp == '6C']
rep1_6C_end <- stratification$DAS35[stratification$rep == 'rep1' & stratification$temp == '6C']
rep1_16C_start <- stratification$DAS14[stratification$rep == 'rep1' & stratification$temp == '16C']
rep1_16C_end <- stratification$DAS35[stratification$rep == 'rep1' & stratification$temp == '16C']

rep2_6C_start <- stratification$DAS14[stratification$rep == 'rep2' & stratification$temp == '6C']
rep2_6C_end <- stratification$DAS35[stratification$rep == 'rep2' & stratification$temp == '6C']
rep2_16C_start <- stratification$DAS14[stratification$rep == 'rep2' & stratification$temp == '16C']
rep2_16C_end <- stratification$DAS35[stratification$rep == 'rep2' & stratification$temp == '16C']

rep3_6C_start <- stratification$DAS14[stratification$rep == 'rep3' & stratification$temp == '6C']
rep3_6C_end <- stratification$DAS35[stratification$rep == 'rep3' & stratification$temp == '6C']
rep3_16C_start <- stratification$DAS14[stratification$rep == 'rep3' & stratification$temp == '16C']
rep3_16C_end <- stratification$DAS35[stratification$rep == 'rep3' & stratification$temp == '16C']

# annotate experiment
rep2_3.slim$experiment <- unlist(mapply(rep2_3.slim$date, rep2_3.slim$hour, FUN = getExperiment))

# are there measurements without annotated experiment or annotated as rep1
any(is.na(rep2_3.slim$experiment))
any(grepl('rep1', rep2_3.slim$experiment))

# annotate replicate and temperature
rep2_3.slim$replicate <- unlist(lapply(rep2_3.slim$experiment, function(exp){strsplit(exp, split = '_')[[1]][2]}))
rep2_3.slim$temperature <- unlist(lapply(rep2_3.slim$experiment, function(exp){strsplit(exp, split = '_')[[1]][1]}))

# annotate DAS
rep2_3.slim$DAS_decimal <- unlist(mapply(rep2_3.slim$rep, rep2_3.slim$temp, rep2_3.slim$date, rep2_3.slim$hour, FUN = getDAS.decimal))
rep2_3.slim$DAS <- floor(rep2_3.slim$DAS_decimal)

# annotate hesoIDs and position 
heso.rep2_3.correct$pot <- paste(heso.rep2_3.correct$tray, heso.rep2_3.correct$coordinate, sep = '_')
rep2_3.slim$pot <- heso.rep2_3.correct$pot[match(rep2_3.slim$Snapshot.ID.Tag, heso.rep2_3.correct$hesoID)]
rep2_3.slim$tray <- unlist(lapply(rep2_3.slim$pot, function(pot){strsplit(pot, split = '_')[[1]][1]}))
rep2_3.slim$coordinate <- unlist(lapply(rep2_3.slim$pot, function(pot){strsplit(pot, split = '_')[[1]][2]}))

# annotate accession
rep2_3.slim$acn <- unlist(mapply(rep2_3.slim$pot, rep2_3.slim$experiment, FUN = getAccession))
```




## Pudding time
The proof of the pudding is in the eating.
To test, heritabilities are calcualte on the reannotated data.
It's expected that heritabilties go up in the range of teh first replicate (around 50%)

```{r heritabilities}
experiments <- unique(rep2_3.slim$experiment)
DASs <- unique(rep2_3.slim$DAS)

h <- data.frame('experiment' = rep(experiments, each = length(DASs)), 'DAS' = rep(DASs, length(experiments)))
h$H <- unlist(mapply(h$experiment, h$DAS, FUN = getH_perExp_perDAS))
h$Ve <- unlist(mapply(h$experiment, h$DAS, FUN = getVe_perExp_perDAS))
h$Vg <- unlist(mapply(h$experiment, h$DAS, FUN = getVg_perExp_perDAS))

pdf('/Volumes/nordborg/user/pieter.clauw/Documents/Experiments/UltimateQandD/Results/Growth/AllReps/Plots/heritabilities_fixedHesoAnnotation.pdf', width = 10)
boxplot(H ~ experiment, data = h)
dev.off()


```




